Supplementary material for ``Screening Sinkhorn Algorithm for regularized Optimal Transport''

### 2019-05-17 tex/v3/section_3_algorithm.tex

\begin{proposition}
\label{prop:bounds_of_usc_and_vsc}
Let $(u^{\text{sc}}, v^{\text{sc}})$ be an optimal solution of problem~\eqref{screen-sinkhorn_second_def}, and $K_{\min} = \min\limits_{i\in I_{\varepsilon,\kappa},j \in J_{\varepsilon,\kappa}}K_{ij}$. Then,
one has
\begin{equation}
\label{bound_on_u}
\frac \varepsilon\kappa \vee \frac{\min_{i \in I_{\varepsilon,\kappa}}\mu_i}{\varepsilon (m- m_b) + \varepsilon \vee \frac{\max_{j\in J_{\varepsilon,\kappa}} \nu_j}{n\varepsilon\kappa K_{\min}} m_b} \leq e^{u^{\text{sc}}_i} \leq \frac \varepsilon\kappa\vee \frac{\max_{i \in I_{\varepsilon,\kappa}} \mu_i}{m\varepsilon K_{\min}},
\end{equation}
and
\begin{equation}
\label{bound_on_v}
\varepsilon\kappa \vee \frac{\min_{j \in J_{\varepsilon,\kappa}}\nu_j}{\varepsilon(n- n_b) + \varepsilon \vee \frac{\kappa\max_{i\in I_{\varepsilon,\kappa}} \mu_i}{m\varepsilon K_{\min} } n_b} \leq e^{v^{\text{sc}}_j} \leq \varepsilon\kappa \vee \frac{\max_{j \in J_{\varepsilon,\kappa}} \nu_j}{n\varepsilon K_{\min} }
\end{equation}
for all $i\in I_{\varepsilon,\kappa}$ and $j\in J_{\varepsilon,\kappa}$.
\end{proposition}

>>>>

\begin{proposition}
\label{prop:bounds_of_usc_and_vsc}
Let $(u^{\text{sc}}, v^{\text{sc}})$ be an optimal solution of problem~\eqref{screen-sinkhorn_second_def} and $K_{\min} = \min\limits_{i\in I_{\varepsilon,\kappa},j \in J_{\varepsilon,\kappa}}K_{ij}$. Then,
one has
\begin{equation}
\label{bound_on_u}
\frac \varepsilon\kappa \vee \frac{\min_{i \in I_{\varepsilon,\kappa}}\mu_i}{\varepsilon (m- m_b) + \varepsilon \vee \frac{\max_{j\in J_{\varepsilon,\kappa}} \nu_j}{n\varepsilon\kappa K_{\min}} m_b} \leq e^{u^{\text{sc}}_i} \leq \frac \varepsilon\kappa\vee \frac{\max_{i \in I_{\varepsilon,\kappa}} \mu_i}{m\varepsilon K_{\min}},
\end{equation}
and
\begin{equation}
\label{bound_on_v}
\varepsilon\kappa \vee \frac{\min_{j \in J_{\varepsilon,\kappa}}\nu_j}{\varepsilon(n- n_b) + \varepsilon \vee \frac{\kappa\max_{i\in I_{\varepsilon,\kappa}} \mu_i}{m\varepsilon K_{\min} } n_b} \leq e^{v^{\text{sc}}_j} \leq \varepsilon\kappa \vee \frac{\max_{j \in J_{\varepsilon,\kappa}} \nu_j}{n\varepsilon K_{\min} }
\end{equation}
for all $i\in I_{\varepsilon,\kappa}$ and $j\in J_{\varepsilon,\kappa}$.
\end{proposition}

### 2019-05-17 tex/v3/supplementary_material.tex

The optimality condition for $({u}^{\text{sc}}, {v}^{\text{sc}})$ entails 
\begin{align}
\label{i-th-marginal-mu} %=({B}({u}^{\text{sc}},{v}^{\text{sc}})\mathbf 1_m)_i
{\mu}^{\text{sc}}_i  &= 
\begin{cases}
e^{u^{\text{sc}}_i} \sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}, \text{ if  }i \in I_{\varepsilon,\kappa},\\
\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}, \text{ if  }i \in I^\complement_{\varepsilon,\kappa}
\end{cases}
=\begin{cases}
\kappa \mu_i, \text{ if  }i \in I_{\varepsilon,\kappa},\\
\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}, \text{ if  }i \in I^\complement_{\varepsilon,\kappa}.
\end{cases}
\end{align}
Using inequality~\eqref{bound_on_v}, we obtain 
\begin{align*}
\norm{\mu^{\text{sc}}}_1 &= \sum_{i \in I_{\varepsilon,\kappa}} \mu^{\text{sc}}_i +  \sum_{i \in I^\complement_{\varepsilon,\kappa}}\mu^{\text{sc}}_i\\
& \overset{\eqref{i-th-marginal-mu}}{=} \kappa \norm{\mu_{I_{\varepsilon,\kappa}}^{\text{sc}}}_1 + \frac \varepsilon\kappa \sum_{i \in I^\complement}\Big( \sum_{j \in J_{\varepsilon,\kappa}} K_{ij} e^{v^{\text{sc}}_j} + \varepsilon\kappa \sum_{j\in J^\complement_{\varepsilon,\kappa}}K_{ij}\Big)\\
& \overset{\eqref{bound_on_v}}{\leq} \kappa \norm{\mu_{I_{\varepsilon,\kappa}}^{\text{sc}}}_1 + (n-n_b) \Big(\frac{m_b\max_{j \in J_{\varepsilon,\kappa}} \nu_j}{n\kappa K_{\min}} + (m-m_b)\varepsilon^2 \Big).
\end{align*}

Similarly, we can prove the same statement for $\norm{\nu^{\text{sc}}}_1$.

@@

Straightforward, 

@@

&= \sum_{i\in I_{\varepsilon,\kappa}} (\kappa-\log(\kappa)-1)\mu_i  + \sum_{i\in I^\complement_{\varepsilon,\kappa}}\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j} - \mu_i + \mu_i \log\Big(\frac{\mu_i}{\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}}\Big)\\
&\leq  \sum_{i\in I^\complement_{\varepsilon,\kappa}}\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j} - \mu_i + \mu_i \log\Big(\frac{\mu_i}{\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}}\Big)

@@

Using~\eqref{bound_on_v}, we have in one hand 

@@

&\geq m_bK_{\min} \frac{m\varepsilon^2K_{\min}\min_{j \in J_{\varepsilon,\kappa}}\nu_j}{\kappa((n-n_b)m\varepsilon^2K_{\min} + m\varepsilon^2K_{\min} + \kappa\max_{i\in I_{\varepsilon,\kappa}}\mu_i)}\\

@@

&\geq \frac{mm_b\varepsilon^2(K_{\min})^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}{\kappa((n-n_b)m\varepsilon^2K_{\min}+ m\varepsilon^2K_{\min} + \kappa\max_{i\in I_{\varepsilon,\kappa}}\mu_i)}\\

@@

&\geq \frac{mm_b\varepsilon^2K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}{\kappa((n-n_b)m\varepsilon^2K_{\min}+ m\varepsilon^2K_{\min} + \kappa\max_{i\in I_{\varepsilon,\kappa}}\mu_i)}.

@@

\frac{1}{\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}} &\leq \frac{\kappa((n-n_b)m\varepsilon^2K_{\min}+ m\varepsilon^2K_{\min} + \kappa\max_{i\in I_{\varepsilon,\kappa}}\mu_i)}{mm_b\varepsilon^2 K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}\\
&\leq \frac{\kappa(n-n_b+ 1)}{m_bK_{\min}\min_{j \in J_{\varepsilon,\kappa}}\nu_j} + \frac{\kappa^2}{mm_b\varepsilon^2K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}.

@@

&\qquad + \max_{i}\mu_i\log\Big(\frac{\kappa(n-n_b+ 1)\max_{i}\mu_i}{m_bK_{\min}\min_{j \in J_{\varepsilon,\kappa}}\nu_j} + \frac{\kappa^2\max_{i}\mu_i}{mm_b\varepsilon^2 K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}\Big)

@@

&\qquad + \max_{i} \mu_i\log\Big(\frac{\kappa(n-n_b+ 1)\max_{i} \mu_i}{m_bK_{\min}\min_{j \in J_{\varepsilon,\kappa}}\nu_j} + \frac{\kappa^2\max_{i} \mu_i}{mm_b\varepsilon^2 K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}\Big).

@@

\norm{{\mu}^\star -{\mu}^{\text{sc}}}^2_1 \leq & n_b(\kappa-\log(\kappa)-1)\max_{i} \mu_i + 7(n-n_b)\bigg(\frac{m_b\max_{j}\nu_j}{n\kappa\min_{i,j} K_{ij}} + (m- m_b) \varepsilon^2 - \min_{i}\mu_i\\
&+ \max_{i} \mu_i\log\Big(\frac{\kappa(n-n_b+ 1)\max_{i} \mu_i}{m_bK_{\min}\min_{j \in J_{\varepsilon,\kappa}}\nu_j} + \frac{\kappa^2\max_{i} \mu_i}{mm_b\varepsilon^2K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}\Big).%\bigg\}^{1/2}

>>>>

The optimality condition for $({u}^{\text{sc}}, {v}^{\text{sc}})$ entails 
\begin{align}
\label{i-th-marginal-mu} 
{\mu}^{\text{sc}}_i  &= 
\begin{cases}
e^{u^{\text{sc}}_i} \sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}, \text{ if  }i \in I_{\varepsilon,\kappa},\\
\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}, \text{ if  }i \in I^\complement_{\varepsilon,\kappa}
\end{cases}
=\begin{cases}
\kappa \mu_i, \text{ if  }i \in I_{\varepsilon,\kappa},\\
\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}, \text{ if  }i \in I^\complement_{\varepsilon,\kappa},
\end{cases}
\end{align}
and 
\begin{align}
\label{i-th-marginal-nu} 
{\nu}^{\text{sc}}_j  &= 
\begin{cases}
e^{v^{\text{sc}}_j} \sum_{i=1}^n K_{ij} e^{u^{\text{sc}}_i}, \text{ if  }j \in J_{\varepsilon,\kappa},\\
\varepsilon\kappa\sum_{i=1}^n K_{ij} e^{u^{\text{sc}}_i}, \text{ if  }j \in J^\complement_{\varepsilon,\kappa}
\end{cases}
=\begin{cases}
\frac{\nu_j}{\kappa}, \text{ if  }j \in J_{\varepsilon,\kappa},\\
\varepsilon\kappa\sum_{i=1}^n K_{ij} e^{u^{\text{sc}}_i}, \text{ if  }j \in J^\complement_{\varepsilon,\kappa}.
\end{cases}
\end{align}

Using inequality~\eqref{bound_on_v}, we obtain 
\begin{align*}
\norm{\mu^{\text{sc}}}_1 &= \sum_{i \in I_{\varepsilon,\kappa}} \mu^{\text{sc}}_i +  \sum_{i \in I^\complement_{\varepsilon,\kappa}}\mu^{\text{sc}}_i\\
& \overset{\eqref{i-th-marginal-mu}}{=} \kappa \norm{\mu_{I_{\varepsilon,\kappa}}^{\text{sc}}}_1 + \frac \varepsilon\kappa \sum_{i \in I^\complement}\Big( \sum_{j \in J_{\varepsilon,\kappa}} K_{ij} e^{v^{\text{sc}}_j} + \varepsilon\kappa \sum_{j\in J^\complement_{\varepsilon,\kappa}}K_{ij}\Big)\\
& \overset{\eqref{bound_on_v}}{\leq} \kappa \norm{\mu_{I_{\varepsilon,\kappa}}^{\text{sc}}}_1 + (n-n_b) \Big(\frac{m_b\max_{j \in J_{\varepsilon,\kappa}} \nu_j}{n\kappa K_{\min}} + (m-m_b)\varepsilon^2 \Big).
\end{align*}
Again by left-hand-side of inequaltiy~\eqref{bound_on_v}, we arrive at 
\begin{align*}
\norm{\mu^{\text{sc}}}_1 %&= \sum_{i \in I_{\varepsilon,\kappa}} \mu^{\text{sc}}_i +  \sum_{i \in I^\complement_{\varepsilon,\kappa}}\mu^{\text{sc}}_i\\
%& \overset{\eqref{i-th-marginal-mu}}{=} \kappa \norm{\mu_{I_{\varepsilon,\kappa}}^{\text{sc}}}_1 + \frac \varepsilon\kappa \sum_{i \in I^\complement}\Big( \sum_{j \in J_{\varepsilon,\kappa}} K_{ij} e^{v^{\text{sc}}_j} + \varepsilon\kappa \sum_{j\in J^\complement_{\varepsilon,\kappa}}K_{ij}\Big)\\
& \overset{}{\geq} \kappa \norm{\mu_{I_{\varepsilon,\kappa}}^{\text{sc}}}_1 + (n -n_b) \Big(\frac{mm_b\varepsilon^2 K_{\min}^2 \min_{j\in J_{\varepsilon, \kappa}} \nu_j}{ (n-n_b)m\kappa \varepsilon^2 K_{\min} + n_b\kappa^2 \max_{i\in I_{\varepsilon,\kappa}}\mu_i }+ (m-m_b)\varepsilon^2K_{\min}\Big),
\end{align*}
which gives the claimed result.
Similarly, we can prove the same statement for $\norm{\nu^{\text{sc}}}_1$.

@@

By~\eqref{i-th-marginal-mu}, we have

@@

&= \sum_{i\in I_{\varepsilon,\kappa}} (\kappa-\log(\kappa)-1)\mu_i  + \sum_{i\in I^\complement_{\varepsilon,\kappa}}\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j} - \mu_i + \mu_i \log\Big(\frac{\mu_i}{\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}}\Big).
% &\leq  \sum_{i\in I^\complement_{\varepsilon,\kappa}}\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j} - \mu_i + \mu_i \log\Big(\frac{\mu_i}{\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}}\Big)

@@

Now by~\eqref{bound_on_v}, we have in one hand 

@@

&\geq m_bK_{\min} \frac{m\varepsilon^2K_{\min}\min_{j \in J_{\varepsilon,\kappa}}\nu_j}{\kappa((n-n_b)m\varepsilon^2K_{\min} + m\varepsilon^2K_{\min} + n_b\kappa\max_{i\in I_{\varepsilon,\kappa}}\mu_i)}\\

@@

&\geq \frac{mm_b\varepsilon^2(K_{\min})^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}{\kappa((n-n_b)m\varepsilon^2K_{\min}+ m\varepsilon^2K_{\min} + n_b\kappa\max_{i\in I_{\varepsilon,\kappa}}\mu_i)}\\

@@

&\geq \frac{mm_b\varepsilon^2K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}{\kappa((n-n_b)m\varepsilon^2K_{\min}+ m\varepsilon^2K_{\min} + n_b\kappa\max_{i\in I_{\varepsilon,\kappa}}\mu_i)}.

@@

\frac{1}{\frac \varepsilon\kappa\sum_{j=1}^m K_{ij} e^{v^{\text{sc}}_j}} &\leq \frac{\kappa((n-n_b)m\varepsilon^2K_{\min}+ m\varepsilon^2K_{\min} + n_b\kappa\max_{i\in I_{\varepsilon,\kappa}}\mu_i)}{mm_b\varepsilon^2 K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}\\
&\leq \frac{\kappa(n-n_b+ 1)}{m_bK_{\min}\min_{j \in J_{\varepsilon,\kappa}}\nu_j} + \frac{n_b\kappa^2\max_{i\in I_{\varepsilon,\kappa}}\mu_i}{mm_b\varepsilon^2K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}.

@@

&\qquad + \max_{i}\mu_i\log\Big(\frac{\kappa(n-n_b+ 1)\max_{i}\mu_i}{m_bK_{\min}\min_{j \in J_{\varepsilon,\kappa}}\nu_j} + \frac{n_b\kappa^2(\max_{i}\mu_i)^2}{mm_b\varepsilon^2 K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}\Big)

@@

&\qquad + \max_{i} \mu_i\log\Big(\frac{\kappa(n-n_b+ 1)\max_{i} \mu_i}{m_bK_{\min}\min_{j \in J_{\varepsilon,\kappa}}\nu_j} + \frac{n_b\kappa^2(\max_{i} \mu_i)^2}{mm_b\varepsilon^2 K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}\Big).

@@

\norm{{\mu} -{\mu}^{\text{sc}}}^2_1 \leq & n_b(\kappa-\log(\kappa)-1)\max_{i} \mu_i + 7(n-n_b)\bigg(\frac{m_b\max_{j}\nu_j}{n\kappa\min_{i,j} K_{ij}} + (m- m_b) \varepsilon^2 - \min_{i}\mu_i\\
&+ \max_{i} \mu_i\log\Big(\frac{\kappa(n-n_b+ 1)\max_{i} \mu_i}{m_bK_{\min}\min_{j \in J_{\varepsilon,\kappa}}\nu_j} + \frac{n_b\kappa^2(\max_{i} \mu_i)^2}{mm_b\varepsilon^2K_{\min}^2\min_{j \in J_{\varepsilon,\kappa}}\nu_j}\Big).%\bigg\}^{1/2}
